Hybrid platforms and bargaining power

### 2024-04-24 src/paper/application.tex
Minor text improvements

The first set of results in this paper go beyond the aforementined studies in the sense that they illuminate the effects of hybridness as a more abstract phenomenon, and thus also highlight the similarities between the hybrid platform literature and adjacent ones.
Certain structures, such as vertical relationships, or even traditional retail stores share a number of features with platforms (notably, the presence of a dominant, indispensable entity), and can be modeled in much the same way.
Therefore, the methodology and results in this paper are also of interest to researchers studying questions such as retailers having their own private labels \parencite{steiner2004nature} or vertical integration in upstream-downstream markets \parencite{hart1990vertical,aghion2006vertical}.
\textcite{de2005vertical} and \textcite{montez2007downstream}, in which bargaining between the upstream and downstream firms take the center stage, are particularly closely related.
Among empirical studies in related settings, \textcite{ho2017insurer} looks at the effect of insurer competition on health care prices, while \textcite{crawford2018welfare} explores the effects of vertical integration on television markets.

This paper is also related to various other strands of the industrial organization literature.
For one, it belongs to the research agenda on understanding platforms and their role in the economy \parencite[e.g.][]{rochet2003platform,hagiu2004optimal,armstrong2006competition,evans2011platform,lee2014competing}.
Furthermore, it is somewhat adjacent to the literature on the importance of exclusive content \parencite[e.g.][]{hagiu2011exclusivity,lee2013vertical,dou2014sell,weeds2016tv}, with the difference that instead of exclusive content giving an advantage against competing platforms, in this paper, own products provide an advantage over potential entrants.
More generally, many results and concepts of this paper are also applicable outside the context of (multi-sided) platforms.

@@

I make use of a reduced-form, but quite general approach to model production and demand.
First, I assume that all fringe firms and consumers are ex-ante identical.
This implies two things: (1) per-consumer profits are just a function of the number of fringe and platform products, and (2) the platform's and the fringe's profits are linear in the number of consumers.\footnote{
    While normally this would exclude fixed production costs, this is not the case here.
    $\Pi$ only describes profits conditional on $N_C$ and $N_F$ players entering the platform.
    Fixed costs can be incorporated into the first stage of the game, where fringe firms decide whether to enter or not.
}
Therefore, the total profits of the platform and the fringe taken together have a simple functional form $N_C f(N_P, N_F)$.\footnote{
    Based on the same reasoning, it is easy to see that total welfare has a similar form.
}
Formally, this is described by the following assumption.
\begin{assumption}
    \label{ass:identical_fringe}
    The total profits of the platform and the fringe are described by 
    \begin{align*}
        \Pi(N_P, N_F, N_C) = N_C f(N_P, N_F),
    \end{align*}
    where $f: \mathbb{R}^+_0 \times \mathbb{R}^+_0 \to \mathbb{R}^+_0$.
\end{assumption}
Here, $N_F$ is the mass of fringe sellers, $N_C$ is the mass of consumers, and $N_P$ measures the degree to which the platform sells its own products.
For example, in a setting with differentiated products and monopolistic competition, $N_F$ and $N_P$ represent the product variety of the fringe and the platform, respectively.
$N_P = 0$ corresponds to the case when the platform is a pure intermediary, and $N_F = 0$ to the case then the platform is a pure retailer.

>>>>

The first set of results in this paper go beyond the aforementioned studies in the sense that they illuminate the effects of hybridness as a more abstract phenomenon, and thus also highlight the similarities between the hybrid platform literature and adjacent ones.
Certain structures, such as vertical relationships, or even traditional retail stores share a number of features with platforms (notably, the presence of a dominant, indispensable entity), and can be modeled in much the same way.
Therefore, the methodology and results in this paper are also of interest to researchers studying questions such as retailers having their own private labels \parencite{steiner2004nature} or vertical integration in upstream-downstream markets \parencite{hart1990vertical,aghion2006vertical}.
\textcite{de2005vertical} and \textcite{montez2007downstream}, in which bargaining between the upstream and downstream firms take the center stage, are particularly closely related.
Among empirical studies in related settings, \textcite{ho2017insurer} looks at the effect of insurer competition on health care prices, while \textcite{crawford2018welfare} explores the effects of vertical integration on television markets.

This paper is also related to various other strands of the industrial organization literature.
Most generally, it is a contribution to the research agenda on understanding platforms and their role in the economy \parencite[e.g.][]{rochet2003platform,hagiu2004optimal,armstrong2006competition,evans2011platform,lee2014competing}.
Furthermore, it is somewhat adjacent to the literature on the importance of exclusive content \parencite[e.g.][]{hagiu2011exclusivity,lee2013vertical,dou2014sell,weeds2016tv}, with the difference that instead of exclusive content giving an advantage against competing platforms, in this paper, own products provide an advantage over potential entrants.
More generally, many results and concepts of this paper are also applicable outside the context of (multi-sided) platforms.

@@

I make use of a reduced-form, but quite general approach to model production and demand.
First, I assume that all fringe firms and consumers are ex-ante identical.
This implies two things: (1) per-consumer profits are just a function of the number of fringe and platform products, and (2) the platform's and the fringe's profits are linear in the number of consumers.\footnote{
    While normally this would exclude fixed production costs, this is not the case here.
    $\Pi$ only describes profits conditional on $N_C$ and $N_F$ players entering the platform.
    Fixed costs can be incorporated into the first stage of the game, where fringe firms decide whether to enter or not.
}

As a consequence, the total profits of the platform and the fringe taken together have a simple functional form $N_C f(N_P, N_F)$.\footnote{
    Based on the same reasoning, it is easy to see that total welfare has a similar form.
}
Formally, this is described by the following assumption.
\begin{assumption}
    \label{ass:identical_fringe}
    The total profits of the platform and the fringe are described by 
    \begin{align*}
        \Pi(N_P, N_F, N_C) = N_C f(N_P, N_F),
    \end{align*}
    where $f: \mathbb{R}^+_0 \times \mathbb{R}^+_0 \to \mathbb{R}^+_0$.
\end{assumption}
Here, $N_F$ is the mass of fringe sellers, $N_C$ is the mass of consumers, and $N_P$ measures the degree to which the platform sells its own products.
For example, in a setting with differentiated products and monopolistic competition, $N_F$ and $N_P$ represent the product variety of the fringe and the platform, respectively.
$N_P = 0$ corresponds to the case when the platform is a pure intermediary, and $N_F = 0$ to the case then the platform is a pure retailer.
